---
title: "getting started _tiger"
author: "Kayla Buchanan"
date: "2022-10-11"
output: pdf_document
---

## Downloading Data

Waves are here: https://www.pewresearch.org/american-trends-panel-datasets/

Remember the codebook for demographic profile variables: https://www.pewresearch.org/wp-content/uploads/2018/05/Codebook-and-instructions-for-working-with-ATP-data.pdf

## Loading Usual Packages

```{r}
library(tidyverse)
```


## Loading Data

The downloaded data is in .sav format. Use the `haven` package to open it. First install and load the package...

```{r}
#install.packages("haven") # Put a hastag in front of this line after installing
library(haven)
```

Then use `read_sav()` rather than `read_csv()` to load the file.
#Reading in Wave 74 & Wave 92
```{r}
w74 <- read_sav("../../data/ATP W74.sav")

w92 <- read_sav("../../data/ATP W92.sav")

names(w92)


```



```{r}
table(w74$F_GENDER)
```

#Wave 72 (Nesting and Pulling Variables )
```{r}
w74 <- w74 |> select(QKEY, F_CDIVISION,F_EDUCCAT2, F_IDEO, F_GENDER, F_RACETHNMOD, SNSUSE_W74, AVOIDPT1_W74,AVOIDPT2_a_W74,AVOIDPT2_b_W74,AVOIDPT2_c_W74,AVOIDPT2_d_W74,AVOIDPT2_e_W74,CANCUL1_W74,CANCUL2_W74, starts_with("WEIGHT"))

w74 <- w74 |> 
  mutate(AVOIDPT1_W74 = ifelse(SNSUSE_W74==2, 5, AVOIDPT1_W74))

w74 <- w74 |> 
  mutate(AVOIDPT2_a_W74 = ifelse (AVOIDPT1_W74<=2, 4, AVOIDPT2_a_W74))

w74 <- w74 |> 
  mutate(AVOIDPT2_b_W74 = ifelse (AVOIDPT1_W74<=2, 4, AVOIDPT2_b_W74))


w74 <- w74 |> 
  mutate(AVOIDPT2_c_W74 = ifelse (AVOIDPT1_W74<=2, 4, AVOIDPT2_c_W74))

w74 <- w74 |> 
  mutate(AVOIDPT2_d_W74 = ifelse (AVOIDPT1_W74<=2, 4, AVOIDPT2_d_W74))
  
w74 <- w74 |> 
  mutate(AVOIDPT2_e_W74 = ifelse (AVOIDPT1_W74<=2, 4, AVOIDPT2_e_W74))

w74 <- w74 |> 
  mutate(AVOIDPT2_a_W74 = ifelse (AVOIDPT1_W74==5, 5, AVOIDPT2_a_W74))
         
w74 <- w74 |> 
  mutate(AVOIDPT2_b_W74 = ifelse (AVOIDPT1_W74==5, 5, AVOIDPT2_b_W74))

w74 <- w74 |> 
  mutate(AVOIDPT2_c_W74 = ifelse (AVOIDPT1_W74== 5, 5, AVOIDPT2_c_W74))

w74 <- w74 |> 
  mutate(AVOIDPT2_d_W74 = ifelse (AVOIDPT1_W74== 5, 5, AVOIDPT2_d_W74))

w74 <- w74 |> 
  mutate(AVOIDPT2_e_W74 = ifelse (AVOIDPT1_W74== 5, 5, AVOIDPT2_e_W74))
```         
#Wave 92 + New Dataset (Pulling Variables )      
```{r}
w92 <- w92 |> select(QKEY,CRIM_SENT2_W92, starts_with("WEIGHT"))


w74_92 <- left_join(w74, w92) %>% 
  filter(!is.na(WEIGHT_W92))

avoid_labels <- c("Major", "Minor", "Not")

```

#labelling variables 
```{r}

w74_92 <- w74_92 |>
     mutate(education = factor(F_EDUCCAT2,
                             levels = c(1, 2, 3, 4,5,6, 99),
                             labels = c("Less Than High School", "High School Graduate","Some College",
                                        "Associate's Degree","College graduate","Postgraduate", "NA")),
            division = factor(F_CDIVISION,
                           levels = c(9,8,7,6,5,4,3, 2, 1),
                           labels = c("Pacific", 
                                      "Mountain", "West South Central", "East South Central", "South Atlantic", "West North Central", "East North Central", "Middle Atlantic", "New England")),
            ideology = factor(F_IDEO, 
                              levels = c(1, 2, 3, 4,5), 
                              labels = c("Very Conservative", "Conservative","Moderate","Liberal","Very Liberal")), 
              gender = factor(F_GENDER, 
                              levels = c(1, 2, 3, 99), 
                              labels = c("Man", "Woman","Other","Refused to Answer")), 
              race = factor(F_RACETHNMOD, 
                              levels = c(1, 2, 3, 4,5, 99), 
                              labels = c("White", "Black","Hispanic","Other","Asian","Refused")), 
            socialmedia = factor(SNSUSE_W74,
                                 levels = c(1, 2),
                                 labels = c("Uses Social Media", "Does not Use Social Media")), 
            avoid_post = factor(AVOIDPT1_W74,
                                levels = c(1 , 2 , 3 , 4) , 
                                labels = c("Often", "Sometimes","Rarely", "Never")),
            cancul1 = factor(CANCUL1_W74, 
                             levels = c(1,2), 
                             labels = c("Accountable", "Punish")), 
            cancul2 = factor(CANCUL2_W74, 
                             levels = c(1, 2, 3, 4),
                             labels = c("A great deal", "A fair amount", "Not too much", "None at all")),
            avoida = factor(AVOIDPT2_a_W74,
                            levels = c(1, 2,3), 
                            labels = avoid_labels),
            avoidb = factor(AVOIDPT2_b_W74,
                            levels = c(1, 2,3), 
                            labels = avoid_labels), 
            avoidc = factor(AVOIDPT2_c_W74,
                            levels = c(1, 2,3), 
                            labels = avoid_labels),
            avoidd = factor(AVOIDPT2_d_W74,
                            levels = c(1, 2,3), 
                            labels = avoid_labels), 
            avoide = factor(AVOIDPT2_e_W74,
                            levels = c(1, 2,3), 
                            labels = avoid_labels), 
            crimsent = factor(CRIM_SENT2_W92, 
                              levels = c(1,2,3), 
                              labels = c("Too much time", "Too little time", "About the right amount")))

```

#Dataset with Categorical Variables + Tables 
```{r}
w74_92 <- w74_92 |> 
  select(QKEY,education, division, ideology, gender, race, socialmedia, avoid_post, cancul1, cancul2, avoida,avoidb, avoidc, avoidd, avoide, crimsent, WEIGHT_W92) 

mytable <- w74_92 |> 
  filter(cancul1!=99) |> 
  filter(crimsent!=99) |> 
  mutate(cancul1 = as.factor(cancul1),
         crimsent = as.factor(crimsent)) |> 
  crosstab(cancul1, crimsent, WEIGHT_W92, n = FALSE) |> 
  pivot_longer(names_to = "crimsent", values_to = "pct", 2:4) |> 
  mutate(crimsent = factor(crimsent,
                           levels = c("Too much time",
                                      "About the right amount",
                                      "Too little time"))) |> 
  ggplot(aes(x = cancul1, y = pct, fill = crimsent)) + geom_col() +
  geom_text(aes(label = round(pct, 2)),
            position = position_stack(vjust = .5),
            size = 8) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 16),
        legend.text = element_text(size = 16)) 

 w74_92 %>% 
  filter(cancul1!=99) %>% 
  mutate(cancul1 = as.factor(cancul1)) %>% 
  crosstab(gender, cancul1, WEIGHT_W92)
 
  w74_92 %>% 
  filter(crimsent!=99) %>% 
  mutate(crimsent = as.factor(crimsent)) %>% 
  crosstab(gender,crimsent, WEIGHT_W92)
  
   w74_92 %>% 
  filter(cancul1!=99) %>% 
  mutate(cancul1 = as.factor(cancul1)) %>% 
  crosstab(race, cancul1, WEIGHT_W92)
 
  w74_92 %>% 
  filter(crimsent!=99) %>% 
  mutate(crimsent = as.factor(crimsent)) %>% 
  crosstab(race,crimsent, WEIGHT_W92)
  
  
     w74_92 %>% 
  filter(cancul1!=99) %>% 
  mutate(cancul1 = as.factor(cancul1)) %>% 
  crosstab(division, cancul1, WEIGHT_W92)
 
  w74_92 %>% 
  filter(crimsent!=99) %>% 
  mutate(crimsent = as.factor(crimsent)) %>% 
  crosstab(division,crimsent, WEIGHT_W92)
  
     w74_92 %>% 
  filter(cancul1!=99) %>% 
  mutate(cancul1 = as.factor(cancul1)) %>% 
  crosstab(education, cancul1, WEIGHT_W92)
 
  w74_92 %>% 
  filter(crimsent!=99) %>% 
  mutate(crimsent = as.factor(crimsent)) %>% 
  crosstab(education,crimsent, WEIGHT_W92)
  
     w74_92 %>% 
  filter(cancul1!=99) %>% 
  mutate(cancul1 = as.factor(cancul1)) %>% 
  crosstab(ideology, cancul1, WEIGHT_W92)
 
  w74_92 %>% 
  filter(crimsent!=99) %>% 
  mutate(crimsent = as.factor(crimsent)) %>% 
  crosstab(ideology,crimsent, WEIGHT_W92)
  
  
```