---
title: "Introducing Visualizations" 
author: "ML"
date: October 25, 2022
output: html_document
---

## Load Usual Packages

```{r}
library(tidyverse)
library(DT)
```

I think every group has used the `weights` package by now but install it if you don't have it yet.

```{r}
#install.packages("weights")
library(weights)
```

We'll also use the `pollster` package when we start working with Team Tiger's data.

```{r}
install.packages("pollster") # hashtag this line after installing
library(pollster)
```

## Team McCallum

```{r}
mccallum <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci485_f22/master/data/cces_08_20.csv")
```

We want to summarize how much more union households voted for the democratic candidate in 2020 than non-union households. Let's set up the data for what we need.

```{r}
mccallum <- mccallum |> 
  mutate(union_hh_diff = (union_hh_yes - union_hh_no)*100)
```

What is the distribution of the new `union_hh_diff` variable? How would you describe this distribution?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###
```{r}
summary(mccallum$union_hh_diff)
```

The most common visualizations when you have one continuous variable are histograms and density plots. Let's set up what we need.

Since we only have one variable, we will define an X in ggplot's aesthetic mapping but not a Y.

```{r}
mccallum |> 
  ggplot(aes(x = union_hh_diff))
```

This is our base layer. We'll add another layer with a geometric function on top of it. For a histogram we would use `geom_histogram()` and for a density plot we will use `geom_density()`. Let's make a density plot.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
mccallum |> 
  ggplot(aes(x = union_hh_diff)) +
  geom_density()
```

This plot is combining all the data from 2008, 2012, 2016, and 2020. It would be better to have a separate line for each presidential election. We can do that by assigning a color in our aesthetic map.

```{r}
mccallum |> 
  ggplot(aes(x = union_hh_diff,
             color = as.factor(year))) +
  geom_density()
```

If we were going to make the final version of a plot in R, we would want to continue cleaning at this point. For example, we could give titles to the plot and the legend (with the `labs()` function), change the axis labels (also with the `labs()` function), and move the legend (with the `theme()` function.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
mccallum |> 
  ggplot(aes(x = union_hh_diff,
             color = as.factor(year))) +
  geom_density() +
  labs(x = "Percentage Point Difference in Democratic Vote in Union Households",
       y = "",
       color = "Year",
       title = "Democratic Advantage in Union Households in Presidential Elections") +
  theme(legend.position = "bottom")
  
```

For our learning guide, we want to tell a story about states. In this example, we will focus on 2020 only. How do we do that?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
mccallum |> 
  filter(year == 2020)
```

We will make a plot showing the democratic advantage in each state. What would that look like?

Now our plot will have a variable on the X axis and a variable on the Y axis. Which is which?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###
```{r}
mccallum |> 
  filter(year == 2020) |> 
  ggplot(aes(x = state, y = union_hh_diff))
```

There will definitely be some cleanup of the axis labels but we can do that later.

We want a bar plot. There is a `geom_bar()` but that only works with one variable. Instead, we will use `geom_col()` for a column plot.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
mccallum |> 
  filter(year == 2020) |> 
  ggplot(aes(x = state, y = union_hh_diff)) +
  geom_col()
```

How could we improve this?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###
```{r}
mccallum |> 
  filter(year == 2020) |> 
  ggplot(aes(x = reorder(state, union_hh_diff), 
             y = union_hh_diff)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This looks like something we could work with. Let's take this data and pull only what we need into Datawrapper. The `DT` package's datatable function with buttons will be very helpful for this.

```{r}
mccallum |> 
  filter(year == 2020) |> 
  select(state, union_hh_diff) |> 
  datatable(rownames = FALSE,
            extensions = 'Buttons', options = list(
    dom = 'Btip', # see explanation below
    buttons = list(list(extend = 'copy', title = NULL))))
```

## Team Tiger

Let's extend the column plot function to fill in a column with values of another variable. We'll use the Pew data for this example.

```{r}
tiger <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci485_f22/master/data/tiger_w74_92.csv")
```

As a reminder, we are looking at the relationship between beliefs about cancel culture and beliefs about formal punishment. 

The `cancul1` question is: In general, when people publicly call out others on social media for posting content that might be considered offensive, are they more likely to 1: hold people accountable for their actions, or 2: punish people who didn't deserve it.

The `crimsent` question is: Overall, would you say people who are convicted of crimes in this country serve: 1: too much time in prison, 2: too little time in prison, or 3: about the right amount of time in prison.

Let's look at the cross tabulation of these survey questions. We can use the `crosstab()` function from the pollster package for this. Note that this function requires a survey weight; we'll use `weight_w92`.

```{r}
tiger |> 
  crosstab(x = cancul1, 
           y = crimsent, 
           weight = weight_w92)
```

Let's plot this. Do we have an X and a Y variable?

The catch here is that we will need to pull all the `crimsent` category names into one column and all the percentages into another column. The nice function to learn for this is `pivot_longer()`. It requires a statement about what you want the column with the existing names to be called (`names_to`), a statement about what you want the column with the existing values to be called (`values_to`), and a statement about which columns to pivot.

```{r}
tiger |> 
  crosstab(x = cancul1, 
           y = crimsent, 
           weight = weight_w92) |> 
  pivot_longer(names_to = "crimsent", # name of new column with categories
               values_to = "pct", # name of new column with numbers
               2:4) # which column locations to pivot
```

That's a good function to know because you can use it with any dataframe. For a shortcut with the `pollster()` package, you can simply add the `format = 'long'` option to the crosstab function.

```{r}
tiger |> 
  crosstab(x = cancul1, 
           y = crimsent, 
           weight = weight_w92,
           format = "long") 
```

Now what is our X, what is our Y, and what are the categories we will "fill" in each column?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
tiger |> 
  crosstab(x = cancul1, 
           y = crimsent, 
           weight = weight_w92,
           format = "long") |> 
  ggplot(aes(x = cancul1, y = pct, fill = crimsent)) +
  geom_col()
```

Probably makes sense to assert the levels of crimsent:

```{r}
tiger |> 
  crosstab(x = cancul1, 
           y = crimsent, 
           weight = weight_w92,
           format = "long") |> 
  mutate(crimsent = factor(crimsent,
                           levels = c("Too little time",
                                      "About the right amount",
                                      "Too much time"))) |> 
  ggplot(aes(x = cancul1, y = pct, fill = crimsent)) +
  geom_col()
```

Let's add some labels showing the percentages in each category. We do this by adding another layer with the `geom_text()` function:

```{r}
tiger |> 
  crosstab(x = cancul1, 
           y = crimsent, 
           weight = weight_w92,
           format = "long") |> 
  mutate(crimsent = factor(crimsent,
                           levels = c("Too little time",
                                      "About the right amount",
                                      "Too much time"))) |>   
  mutate(pct = round(pct, 1)) |> 
  ggplot(aes(x = cancul1, y = pct, fill = crimsent)) + 
  geom_col() +
  theme(legend.position = "bottom") +
   geom_text(aes(label = pct), position = position_stack(vjust = 0.5))
```

Let's try this one in Datawrapper! As a general rule, Datawrapper prefers data to be in wide rather than long format:

```{r}
tiger |> 
  mutate(crimsent = factor(crimsent,
                           levels = c("Too little time",
                                      "About the right amount",
                                      "Too much time"))) |> 
  crosstab(x = cancul1, 
           y = crimsent, 
           weight = weight_w92) |> 
  select(-n) |> 
  datatable(rownames = FALSE,
            extensions = 'Buttons', options = list(
    dom = 'Btip', # see explanation below
    buttons = list(list(extend = 'copy', title = NULL))))
```

There is also a `crosstab_3way()` function in pollster if you want to examine how an association between x and y varies by values of a control variable (z). In this example, our control variable is `ideology`. 

```{r}
tiger |> 
  mutate(crimsent = factor(crimsent,
                           levels = c("Too little time",
                                      "About the right amount",
                                      "Too much time")),
         ideology = factor(ideology,
                           levels = c("Very Liberal",
                                      "Liberal",
                                      "Moderate",
                                      "Conservative",
                                      "Very Conservative"))) |> 
  crosstab_3way(x = cancul1, 
                y = crimsent, 
                z = ideology, 
                weight = weight_w92) |> 
  select(-n) |> 
  datatable(rownames = FALSE,
            extensions = 'Buttons', options = list(
    dom = 'Btip', # see explanation below
    buttons = list(list(extend = 'copy', title = NULL))))
```


## Team Han

Now let's move from plots with columns to plots with points or dots.

Let's start by looking at the relationship between zip codes' unemployment rates and their eviction rates. What would you expect? What are some possible ways to summarize the association?

```{r}
han <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci485_f22/master/data/han_data_complete_pop.csv")
```

What's the correlation?

```{r}
cor(han$unemploymentrate, han$evict_pct, use = "complete")
```

What's the correlation weighted by population?

```{r}
wtd.cor(han$unemploymentrate, han$evict_pct, han$population)
```

To visualize the association between two continuous variables, good idea to try a scatterplot first. Use the `geom_point()` function for that.

```{r}
han |> 
  ggplot(aes(x = unemploymentrate, y = evict_pct)) + 
  geom_point()
```

Scatterplots often include a "line of best fit" which we can incorporate by adding a layer with a `geom_smooth()` function where the points are "smoothed" according to the linear model (lm) method.

```{r}
han |> 
  ggplot(aes(x = unemploymentrate, y = evict_pct)) + 
  geom_point() +
  geom_smooth(method = "lm") # for line of best fit
```

(NOTE: Need to clarify what the `evict_pct` values are.)

This plot has points for zip codes across nine cities. What have we seen so far to be able to identify individual cities?

Colors could work. Facets could work too...but watch the scales!

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
han |> 
  ggplot(aes(x = unemploymentrate, y = evict_pct, 
             color = xsite)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_wrap(~xsite, scales = "free") + guides(color = "none")
```

We won't make the scatterplot in Datawrapper. Instead, we will extend the `geom_point()` function to make a dot plot summarizing the eviction rate by the majority race in zip codes for each city. Let's get those eviction rates.

```{r}
han |> 
  group_by(xsite, majority_race) |> 
  summarise(mean_evict_pct = wtd.mean(evict_pct, weights = population))
```

Now we can build a plot. What's our x? What's our y?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
han |> 
  group_by(xsite, majority_race) |> 
  summarise(mean_evict_pct = wtd.mean(evict_pct, weights = population)) |> 
  ggplot(aes(x = mean_evict_pct, y = xsite, color = majority_race)) + geom_point() + theme(legend.position = "bottom")
```

This will look a lot better in Datawrapper. Before creating it there, we'll need the data in wide format rather than long format. This time we can use the `pivot_wider()` function:


```{r}
han |> 
  group_by(xsite, majority_race) |> 
  summarise(mean_evict_pct = wtd.mean(evict_pct, weights = population)) |> 
  mutate(xsite = str_to_title(xsite),
         majority_race = str_to_title(majority_race),
         mean_evict_pct = mean_evict_pct * 100) |> 
  pivot_wider(names_from = "majority_race", # which variable's values will become the new column names
              values_from = "mean_evict_pct") |> # which variable's values will become the values of the new columns
  datatable(escape = TRUE,
            rownames = FALSE,
            extensions = 'Buttons', options = list(
    dom = 'Btip', # see explanation below
    buttons = list(list(extend = 'copy', title = NULL))))
```



