---
title: "Zip Code Characteristics Explaining Race-Eviction Association"
author: "ML"
date: "2022-11-11"
output: html_document
---

Load packages and data
```{r}
library(tidyverse)
library(weights)


han <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci485_f22/master/data/han_data_complete_pop.csv")

# Drop Phoenix
han <- filter(han, xsite != "PHOENIX")
```


# Baseline model

Start with a model regressing eviction on the nonwhite percentage. Use the standardized values.

```{r}
baseline <- lm(scale(evict_pct) ~ scale(nonwhite_pct) + factor(xsite), data = han, weight = population)

summary(baseline)
```

Save the nonwhite_pct coefficient from the baseline 

```{r}
baseline_coef = coef(baseline)[[2]]
```


# Model formula setup

Create vector of the names of the other zip code characteristics we will control for

```{r}
vars <- c("median_hh_income",
            "gini",
            "prop_owner_occ",
            "prop_renter_occ",
            "unemploymentrate",
            "singlerate",
            "rent_pct_incomeE",
            "snap_pct",
            "Less_than_15_minutes",
            "college_plus")
```


Now we'll start building each of the formulas. Start with what is constant in each formula before adding the control variables.

```{r}
nocontrols <- paste("scale(evict_pct)", 
                    "~ scale(nonwhite_pct) + ")
```

Across each of the values of vars, we are going to create a function. That function will paste the chunk above, insert one of the values of vars, and paste the rest of the model. 

```{r}
nocontrols_vars <- 
  unlist(lapply(vars, function(x) paste0(nocontrols, 
                                           "scale(", x, ") + factor(xsite)")))
```

Return `nocontrols_vars` to confirm that we have all ten models

```{r}
nocontrols_vars
```

Now we have to assert that each of the models is a formula

```{r}
formulas <- 
  lapply(nocontrols_vars, formula)
```

Finally we can run each formula as a regression. Note that we are adding the data and weight here rather than earlier.

We are going to pull the second coefficient from the output for each regression model. That is the coefficient for nonwhite_pct controlling for the zip code characteristic in each model and the city. 

```{r}
  models <- lapply(formulas, FUN=function(x) {
    coef(lm(x, data = han, weight = population))[[2]]
  })
```

The function output is a list. We can recover any coefficient by identifying its location in the list. For example, for the first model:

```{r}
models[[1]]
```

If we did everything correctly, that coefficient should be the coefficient for nonwhite_pct in the model controlling for median_hh_income (the first variable in the vector of variables):

```{r}
summary(lm(scale(evict_pct) ~ scale(nonwhite_pct) + scale(median_hh_income) + factor(xsite), data = han, weight = population))
```

Looks good!

Use `unlist` to pull the coefficients from all the models into a vector.

```{r}
coefficients <- unlist(models)
```


Put everything together in a new dataframe.

```{r}
results <- bind_cols( # bind vectors as columns
  vars, # vector of variable names
  coefficients) |> # vector of coefficients unlisted in chunk above
  rename(variable = 1, 
         coefficient = 2)

results
```

Last step is to calculate the proportional difference between the coefficient for nonwhite_pct in the baseline model and the coefficient for nonwhite_pct in each model controlling for a different characteristic.

```{r}
results <- results |> 
  mutate(explained =
           (baseline_coef - coefficient) / 
           baseline_coef)

results
```


